\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-05-10 paper/paper.tex
Revert "Modified model and priors sections"
This reverts commit 04f131ee55044338130f0925a8c7d74ee45d273f.

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Failure data for series systems may be collected at the system level by end-users interested only in the lifetime of the system. In the case of complex, high-reliability systems where cause of failure is not available, a model of failure which
\begin{enumerate}[a]
\item can be estimated from the observed data
\item is flexible enough to fit the observed data
\item relies on an approximation of the understood underlying process
\end{enumerate}
is desireable. We present a Bayesian, hierarchical modeling approach for grouped failure data. We demonstrate this approach with a large, heavily censored real data set with multiple right-censoring and left truncation times. We extend the generalized limited failure population model (GLFP) by \citet{chan}, which provides a generative model for failure times in a population with infant mortality. This model is an example of a bathtub distribution, exhibiting distinct early and late failure modes. The GLFP model assumes there are two causes of failure, one of which affects only a fraction of the population, and that the cause of failure is unobserved. We extend this model to grouped data and show that by borrowing information across groups we can acheieve good model fits for almost all groups. By working with posterior samples, we can easily make inference on a wide range of quantities of interest while accounting for uncertainty.

\subsection{Background}
Often products can fail due to more than one failure mode.  For example, there are many parts in a washing machine that can break causing the whole machine to fail.  Decorative tree lighting can fail if one bulb in the chain burns out.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \cite{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  The GFLP model is a special case of the $s$ independent competing risks model with 2 modes of failure: defects and wearout.  The GFLP model was first introduced to describe the early and late failure modes of CB radios \cite{chan}.  Chan and Meeker used maximum likelihood (ML) estimation to estimate the model parameters, and developed likelihood based approximation methods to construct confidence intervals for functions of model parameters.  Other authors, for example, Basu et al., have employed a Bayesian approach to fit the GFLP model to masked failure data from an engineering system \cite{basu}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  There has not been, however, any extensions of the GFLP model to a hierarchical framework.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.  There are numerous features of the data, however, that make standard estimation approaches for this parametric model problematic.  The first issue is many hard drives have been running for quite some time before entering the sample.  Even hard drives starting after 2013 appear in the data after hundreds of hours of operation.  When an observation is at risk before entering the study, this is known as left truncation.  Also, although we have four years of data, hard drives rarely fail so many units are still in service.  Thus, we have both left truncation and right censoring.  The other challenge is each hard drive model varies in number units in operation and total time on test.  Some models have hundreds of units in service; others have as few as two.  Combined with the truncation and censoring this makes stable ML estimates difficult to obtain for many drive models. \\

Another issue is product populations often contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are unknown, which presents challenges for parameter identification in the GFLP model, as discussed in Chan and Meeker (1999).  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  This is especially helpful for the Backblaze data as there are many drive models with one or two failures.

>>>>

\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Failure data for series systems may be collected at the system level by end-users interested only in the lifetime of the system. In the case of complex, high-reliability systems where cause of failure is not available, a model of failure which
\begin{enumerate}[a]
\item can be estimated from the observed data
\item is flexible enough to fit the observed data
\item relies on an approximation of the understood underlying process
\end{enumerate}
is desirable. We present a Bayesian, hierarchical modeling approach for grouped failure data. We demonstrate this approach with a large, heavily censored real data set with multiple right-censoring and left truncation times. We extend the generalized limited failure population model (GLFP) by \citet{chan}, which provides a generative model for failure times in a population with infant mortality. This model is an example of a bathtub distribution, exhibiting distinct early and late failure modes. The GLFP model assumes there are two causes of failure, one of which affects only a fraction of the population, and that the cause of failure is unobserved. We extend this model to grouped data and show that by borrowing information across groups we can achieve good model fits for almost all groups. By working with posterior samples, we can easily make inference on a wide range of quantities of interest while accounting for uncertainty.

\subsection{Background}
In engineering applications, a product can often fail due to one out of a set of possible malfunctioning components.  For example, a computer system can fail if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of normal wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across $s$ different components or risks \citet{nelson}.  A common assumption in series systems is the time to failure for each risk, $s$, is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all $s$ risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  With engineering systems data, however, the exact cause of failure is frequently unknown or masked from the researcher.  \\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed the cause of failure for each observation came from a known subset of failure modes, and estimation was performed using a Bayesian approach \cite{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et. al used a Bayesian analysis with informative priors to better identify early versus late failure modes without making any data assumptions \cite{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \cite{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \cite{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization algorithm with ML was used, in addition to Bayesian estimation.

\subsection{Motivation}
The goal of this paper is to model, and compare, the lifetime distribution of the various hard drive brands using the GFLP lifetime model.  Product populations often contain a mixture of defective and nondefective units.  The hazard function for this type of population is often described as a bathtub curve: the beginning of the curve corresponds to defective units failing early, followed by a constant hazard, and then and upswing as units fail from wearout.  Ignoring this hazard structure, which is known to exist for computer disk drives, could lead to spurious inference when comparing the reliability of hard drive brand models \cite{chan}.  The general limited failure population model (GFLP) combines different failure time distributions to account for bending hazard functions; for example, one parametric model for defective units (infant mortality) and a second model for wearout \cite{chan}.  Unfortunately, as often the case with lifetime data, the true causes of failure for the hard drives are masked, which presents challenges for parameter identification in the GFLP model.  Therefore, we propose to fit the GFLP model using a hierarchical Bayesian approach.  With over 60 brand drive models in testing a hierarchical framework is advantageous as it pools information from across brands to get more precise estimates of the individual parameters.  In cases where drive models with only one or two failures the hierarchical model shrinks unstable parameter estimates back towards the overall group average.  This is advantageous since for many groups it was not possible to estimate the GFLP model individually by drive model.
